### 2025-05-09 appendices/arima2/appendix.Rnw
Worked on appendices, table, arima2 chapter

\chapter{Appendix for Chapter 2}
\label{chpt:appendix_arima}

<<packages, include=FALSE, echo=FALSE>>=
root <- ""
library(tidyverse)
library(arima2)

myround <- function(x, digits = 1) {
  # taken from the broman package
  if (digits < 1)
    stop("This is intended for the case digits >= 1.")
  if (length(digits) > 1) {
    digits <- digits[1]
    warning("Using only digits[1]")
  }
  tmp <- sprintf(paste("%.", digits, "f", sep = ""), x)
  zero <- paste0("0.", paste(rep("0", digits), collapse = ""))
  tmp[tmp == paste0("-", zero)] <- zero
  tmp
}

theme_set(theme_bw() + theme(plot.title = element_text(hjust = 0.5)))
@

%%%%%%%%%% START

\section{Comparing ARIMA algorithm to Python software}
\label{sec:appendix_arima_python}

<<loadPythonRes, echo=FALSE>>=
python_MLEs <- read.csv(paste0(root, 'data/pythonResults.csv'))
sub_fits <- read.csv(paste0(root, 'data/sub_fits.csv'))[, -1]
best_lls <- readRDS(paste0(root, 'data/sub_fits_lls.rds'))
num_improve <- sum(best_lls - python_MLEs$loglik > 0)
num_better <- sum(best_lls - python_MLEs$loglik > 0.01)

diffs_df <- data.frame(
  model = 1:100,
  diff = best_lls - python_MLEs$loglik
)
@

\noindent Although Algorithm~\ref{alg:mle} is language agnostic, our implementation is in R.
Consequently, all simulation studies for this article use R's \code{stats::arima} function for baseline comparisons.
To demonstrate applicability to other software environments, we briefly compare model likelihoods fit using Python's \code{statsmodels.tsa} module against our implementation of Algorithm~1.
We generate 100 unique Gaussian ARMA$(2, 1)$ models and datasets (each with $n = 100$) where R's \code{stats::arima} provides sub-optimal estimates. These observations are used to fit ARMA$(2, 1)$ models in Python.
Despite these datasets being chosen for sub-optimal results in R, the log-likelihoods in both R and Python are roughly equivalent---a result of both software packages using the same general approach to fitting model parameters.

Our implementation of Algorithm~1 resulted in higher log-likelihoods for $\Sexpr{num_improve}$ out of the 100 datasets compared to Python.
The log-likelihood deficiencies for the remaining three datasets were all smaller than $\epsilon = 10^{-5}$, which is smaller than the tolerance level to be considered as an improvement in our other simulation studies.
While potentially insignificant, these differences can be eliminated by increasing the number of parameter initializations and the convergence criteria of our algorithm.
Directly implementing our algorithm in Python would further eliminate the possibility of these discrepancies entirely.

\section{Uniform Sampling}
\label{sec:appendix_arima_sample}

A common approach to optimizing a non-convex loss function is to perform the optimization routine with distinct parameter initializations.
For ARMA models, picking suitable initialization is a challenging problem that we address with Algorithm~1.
An alternative approach would involve sampling each coefficient independently.
To see why an independent sampling scheme is not used, consider an AR(2) model.
An initialization with parameters $(\phi_1, \phi_2) = (1.1, 0.1)$ is not a valid initialization---as the polynomial roots lie outside the complex unit circle---whereas $(\phi_1, \phi_2) = (1.1, -0.2)$ is perfectly acceptable.
Our algorithm accounts for the complex relationship between model parameters when obtaining random initializations.

To visualize why an independent sampling scheme is not used, consider sampling parameters from a Uniform$(-1, 1)$ distribution.
In Fig~\ref{fig:AR2}, we plot inverted roots that are a result of sampling from this distribution for AR(2) and AR(3) models.
The figure illustrates that a significant percentage of uniformly sampled parameter initializations lie outside the accepted region, and, critically, the entire region of possible initializations is not well covered by uniform sampling.
Picking a uniform distribution with different bounds---or any other independent sampling distribution---results in similar problems.
In order to uniformly sample from the possibly regions, it is necessary to account for the geometry of parameter space, a problem solved by Algorithm~\ref{alg:mle}.

\begin{figure}[ht]
<<ARrootsUnif, fig.height=2.3, echo=FALSE>>=
N <- 1000

coefs_AR2  <- runif(N*2, -1, 1) |> matrix(nrow = N, ncol = 2)

ARs_AR2  <- coefs_AR2[, 1:2]
# MAs <- coefs[, 4:6]

colnames(ARs_AR2)  <- c("ar1", "ar2")

ar_roots_AR2 <- t(apply(ARs_AR2, 1, arima2::ARMApolyroots))
inv_ar_roots_AR2 <- 1/ar_roots_AR2

complex_numbers_AR2  <- as.vector(inv_ar_roots_AR2)

# Extract real and imaginary parts
real_parts_AR2 <- Re(complex_numbers_AR2)
imag_parts_AR2 <- Im(complex_numbers_AR2)

df_AR2 <- data.frame(
  Re = real_parts_AR2,
  Im = imag_parts_AR2,
  alg = 'AR(2) Roots'
)

coefs_AR3 <- runif(N*3, -1, 1) |> matrix(nrow = N, ncol = 3)

ARs_AR3  <- coefs_AR3[, 1:3]

colnames(ARs_AR3)  <- c("ar1", "ar2", "ar3")

ar_roots_AR3 <- t(apply(ARs_AR3, 1, arima2::ARMApolyroots))
inv_ar_roots_AR3 <- 1/ar_roots_AR3

complex_numbers_AR3  <- as.vector(inv_ar_roots_AR3)
real_parts_AR3 <- Re(complex_numbers_AR3)
imag_parts_AR3 <- Im(complex_numbers_AR3)

df_AR3 <- data.frame(
  Re = real_parts_AR3,
  Im = imag_parts_AR3,
  alg = "AR(3) Roots"
)

df_combined <- bind_rows(df_AR2, df_AR3)

ggplot(df_combined, aes(x = Re, y = Im, col = Re^2 + Im^2 >= 1, shape = Re^2 + Im^2 >= 1)) +
  geom_point() +
  theme_bw() +
  annotate("path", x = cos(seq(0, 2*pi, length.out = 100)),
           y = sin(seq(0, 2*pi, length.out = 100))) +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0) +
  coord_fixed() +
  facet_wrap(~alg) +
  scale_color_manual(values = c("black", 'red')) +
  scale_shape_manual(values = c(20, 4)) +
  theme(legend.position = 'none', axis.title = element_blank())
@
\caption{\label{fig:AR2}Inverted roots of 1000 samples of AR(2) and AR(3) coefficients sampled independently from a U$(-1, 1)$ distribution. The red ``x"s are points that lie outside the accepted region, which represents $\Sexpr{round(100*sum((real_parts_AR2^2 + imag_parts_AR2^2 >= 1)) / length(real_parts_AR2), 1)}\%$ of the AR(2) coefficients and $\Sexpr{round(100*sum((real_parts_AR3^2 + imag_parts_AR3^2 >= 1)) / length(real_parts_AR3), 1)}\%$ of the AR(3) coefficients. Increasing the width of the uniform sampling distribution results in a larger fraction outside the accepted region, and decreasing the width results in worse coverage of the range of acceptable parameter values.}
\end{figure}

%%%%%%%%%% END