### 2025-05-11 front_materials/acknowledgments.tex
Finished rough-draft of introduction

I am deeply grateful to all of the mentors in my life who have helped me get to this point.
I am particularly grateful for my advisor, Dr. Edward L. Ionides, who has provided invaluable support for me throughout my academic journey and has helped shape my beliefs about statistics and higher education.
I wish to also thank the other members of my committee who have provided their advise and mentorship. 
Dr. Aaron A. King

Most importantly I would like to thank my spouse, Haylee Wheeler, who has been incredibly supportive and encouranging though my entire academic career.
The progress that I have made as a scholar would not have been possible without her.
You have always been willing to listen to my complaints and struggles, and been with me to celebrate life's victories.