### 2025-05-12 front_materials/acknowledgments.tex

I am deeply grateful to all of the mentors in my life who have helped me get to this point.
I am particularly grateful for my advisor, Dr. Edward L. Ionides, who has provided invaluable support for me throughout my academic journey and has helped shape my beliefs about statistics and higher education.
You have been an incredible mentor to me over the years.

I wish to also thank the other members of my committee who have provided their advise and mentorship. 
Dr. Aaron A. King

I have also been fortunate to have a great group of friends and classmates that have supported me throught my PhD studies.

Most importantly I would like to thank my spouse, Haylee Wheeler, who has been incredibly supportive and encouranging though my entire academic career.
The progress that I have made as a scholar would not have been possible without her.
You have always been willing to listen to my complaints and struggles, and been with me to celebrate life's victories.

This dissertation contains three related projects described in Chapters~\ref{chpt:arima}, \ref{chpt:haiti}, and \ref{chpt:mpif}.
The projects outlined in each of the chapters have either been published in a peer-reviewed journal (Chapter~\ref{chpt:haiti}), or are currently being prepared for publication (Chapters~\ref{chpt:arima} and \ref{chpt:mpif}). 
The projects include work that was done in conjunction with the coauthors of these papers, and I would like to thank them for their work and highlight their contributions.
\begin{itemize}
  \item  In Chapter~\ref{chpt:arima}, my advisor Edward L. Ionides met with me weekly to discuss ideas and progress related to this project.
  He also provided valuable feedback and suggestions on how to present my algorithm and results, and was instrumental in refining the manuscript text. 
  Importantly, he was the first to notice the apparent optimization issue present in existing algorithms, which was the original motivation for this work.
  \item Chapter~\ref{chpt:haiti} was a large undertaking by several members of Dr. Ionides's research group, and I had the privilege of supervising a group of undergraduate students (AnnaElaine Rosengart, Kevin Tan, and Noah Treutle) as well as a master student (Zhuoxun (Josh) Jiang) who worked on this project. 
  All coauthors were part of weekly discussions on the project and helped refine the manuscript text.
  AnnaElaine had a major role in writing software to build Model~1. 
  Similarly, Josh wrote code that enabled us to reproduce existing results that were obtained using Model~2.
  Kevin Tan and Noah Treutle begin investigations related to the linear reductions in transmission rate for Model~1.
  This work is now published in PLoS Computational Biology \citep{wheeler24}. 
  \item Chapter~\ref{chpt:mpif} represents joint work with Dr. Ionides and Aaron Abkemeier, a classmate and coauthor of this project. 
  Aaron is an expert in Measels transimssion models, and is primarily responsible for the analysis presented in Section~\ref{} TODO. 
\end{itemize}

TODO:  
I would also like to acknowledge funding opportunities that have helped make this work possible.
RMF, Rackham graduate grant, Ed's grant, others?