### 2025-05-16 appendices/mpif/assumptions.tex
Addressed pre-defense review comment: Text on page 112 exceeds past the 1" right side margin

\section{Assumptions for Theorem~1}\label{sec:assumptions}

Let $R$ be a separable Banach space, and $\mathcal{B}(R)$ be the Borel $\sigma$-algebra on $R$. We assume that $\Theta \in \mathcal{B}(R)$ is a regular compact set, following Definition~1 of \citet{chen24}), and that $\mathcal{B}(\Theta)$ is the Borel $\sigma$-algebra on $\Theta$.
For all constants $(\epsilon, r) \in (0, \infty) \times \mathbb{R}$, define $B_\epsilon(r) = \{r' \in \mathbb{R}: \|r - r'\| < \epsilon\}$.
Furthermore, if $\{\mu_n\}_{n \geq 1}$ is a sequence of random probability measures on $\big(R, \mathcal{B}(R)\big)$ with $\mathcal{F}_n$ denoting the corresponding filtration, then we denote $K_{\mu_n}$ to be the Markov kernel such that $\theta \sim K_{\mu_n}(\theta', d\theta) \iff \theta \overset{dist}{=}\theta' + U,$ where $U|\mathcal{F}_n \sim \mu_n$.

Let $S_m = \sum_{u = 1}^m N_u + 1$. 
We define the Markov kernel as a sequence in $n' \in \mathbb{N}$, such that $K_{n'}(\theta_{n'-1}, d\theta_{n'}) = h_{u, n}(\theta_{n'}|\theta_{n'-1};\sigma)d\theta_{n'}$ whenever $(m-1)S_U + S_{u-1} < n' \leq (m-1)S_U + S_{u}$.
The following assumption provide sufficient conditions for Theorem~\ref{theorem:pif}.

  \begin{enumerate}
    \item $L(\theta; \mathbf{y}^*) > 0$ for all $\theta \in \Theta$ and $\sup_{(\theta, x_{u, n})}f_{Y_{u, n}|X_{u, n}}(y^*_{u, n}|x_{u,n}; \theta) < \infty$ for all $u \in \seq{1}{U}$ and $n \in \seq{1}{N_u}$. \label{assumption:1}
    \item The transition and measurement densities are sufficiently smooth functions of $\theta$, in the sense that for any $\theta, \theta' \in \Theta$, there exists a a continuous and strictly increasing function $g: [0, \infty) \rightarrow [0, \infty)$ and sequence of measurable functions $\varphi_{u, 0:N_u}: \mathcal{X}^2 \rightarrow \mathbb{R}$, that satisfy:
%     \begin{align}
% \big\vert \log\big( & f_{Y_{u, n}|X_{u, n}}(y^*_{u, n}|x_{u, n};\theta)
% f_{X_{u, n}|X_{u, n-1}}(x_{u, n}|x_{u, n-1};\theta)\big) \notag \\
% & - \log\big(f_{Y_{u, n}|X_{u, n}}(y^*_{u, n}|x_{u, n};\theta')
% f_{X_{u, n}|X_{u, n-1}}(x_{u, n}|x_{u, n-1};\theta')\big) \big\vert \notag \\
% & \leq g\big(\|\theta - \theta'\|\big)\varphi_{u, n}(x_{u, n-1}, x_{u, n}),
% \end{align}
    \begin{align*}
    \big\vert\log\big( & f_{Y_{u, n}|X_{u, n}}(y^*_{u, n}|x_{u, n};\theta)f_{X_{u, n}|X_{u, n-1}}(x_{u, n}|x_{u, n-1};\theta)\big) \\
    & - \log\big(f_{Y_{u, n}|X_{u, n}}(y^*_{u, n}|x_{u, n};\theta')f_{X_{u, n}|X_{u, n-1}}(x_{u, n}|x_{u, n-1};\theta')\big)\big\vert \\
    & \qquad\quad \leq g\big(\|\theta - \theta'\|\big)\varphi_{u, n}(x_{u, n-1}, x_{u, n}),
    \end{align*}
    With the constraint that for all $u\in \seq{1}{U}$, there exists a $\delta_u \in (0, \infty)$ such that
    \begin{align*}
    &  \int \exp\Big\{\delta_u \sum_{n = 0}^{N_u}\varphi_{u, n}(x_{u, n-1}, x_{u, n})\Big\} f_{X_{u, 0}}(x_{u, 0};\theta) \times
    \\
    & \hspace{20mm} \prod_{n = 1}^{N_u}f_{Y_{u, n}|X_{u, n}}(y^*_{u, n}|x_{u, n};\theta)f_{X_{u, n}|X_{u, n-1}}(x_{u, n}|x_{u, n-1};\theta)\, dx_{u, 0:N_u} < \infty,
    \end{align*}
    using the convention that if $n=0$, then
    \begin{eqnarray*}
      \varphi_{u, n}(x_{u, n-1}, x_{u, n}) &=& \varphi_{u, 0}(x_{u, 0}, x_{u, 0})
      \\
      f_{Y_{u, n}|X_{u, n}}(\cdot | x_{u, n}; \theta) &=& 1
      \\
      f_{X_{u, n} | X_{u, n-1}}(x_{u, n}|x_{u, n-1};\theta) &=& f_{X_{u, 0}}(x_{u, 0}; \theta)
    \end{eqnarray*}
    \label{assumption:2}

    \item Let $\{U_{n'}\}_{n' \geq 1}$ be a sequence of $\Theta$-valued random variables such that for all $n' \geq 1$ and sets $A_1, \ldots, A_{n'} \in \mathcal{B}(\Theta)$, 
    $$
    \prob\big(U_k \in A_k, k\in \{1, \ldots, n'\} | \mathcal{F}_{n'}\big) = \prod_{k = 1}^{n'} \mu_k(A_k). 
    $$
    Then the sequence of probability measures $\{\mu_{n'}\}_{n' \geq 0}$ satisfy \label{assumption:kernel}
    \begin{enumerate}
      \item $\mu_0\big(B_{\epsilon}(\theta)\big) > 0$ for all $\theta \in \Theta$ and $\epsilon \in (0, \infty)$. 
      \item There exists a family of $(0, 1]$-valued random variables $(\Gamma^\mu_\delta)_{\delta \in (0, \infty)}$ such that, for all $\delta \in (0, \infty)$, we have $\prob\big(\inf_{n' \geq 1}\mu_{n'}\big(B_\delta(0)\big) \geq \Gamma_{\delta}^\mu\big) = 1$.
      \item $\prob\big(\inf_{n' \geq 1}\inf_{\theta' \in \Theta}\int_\Theta K_{\mu_{n'}}(\theta', d\theta) \geq \Gamma^\mu\big) = 1$ for some $(0, 1]$-valued random variable $\Gamma^\mu$.
      \item There exists a sequence of natural numbers $\{k_{n'}\}_{n' \geq 1}$ and, for all $l \in \mathbb{N}_0$, a sequence of $(0, \infty]$ valued functions $\{g_{n', l}\}_{n' \geq 1}$ defined on $(0, \infty)$ such that \begin{enumerate} 
  \item $\lim_{n \rightarrow \infty} k_{n'} / {n'} = \lim_{n' \rightarrow \infty} 1/k_{n'} = 0$, and for all $\epsilon \in (0, \infty)$, $\lim_{n' \rightarrow \infty}g_{n', l}(\epsilon) = 0$. 
  \item For all $n \geq 1$ such that $n > 2k_{n'}$, $\kappa_{n'} \in \{k_{n'}, \ldots, 2k_{n'}\}$, and $\epsilon \in (0, \infty)$,
  $$
  \frac{1}{t - \kappa_{n'}}\log \prob \big(\exists s \in \{\kappa_{n'} + 1, \ldots, n'\}: \sum_{i = \kappa_{n'} + 1}^s U_i \notin B_{\epsilon}(0) | \mathcal{F}_{n'}\big) \leq -\frac{1}{g_{n', l}(\epsilon)}.
  $$
  \item For all $n' > l$ and $\epsilon \in (0, \infty)$, we have 
  $$
  \frac{1}{n'-l}\log \prob \big(\sum_{i = l + 1}^{s} U_i \in B_{\epsilon}(0), \forall s \in \{l + 1, \ldots, n'\}|\mathcal{F}_{n'}\big) \geq -g_{n', l}(\epsilon).
  $$
\end{enumerate}

    \end{enumerate}
  \end{enumerate}