### 2025-05-16 chapters/introduction/intro.Rnw
Addressing Ed's comments

\chapter{Introduction}
\label{chpt:intro}

The importance of time series analysis has continued to grow with the demand for both the collection and analysis of data.
Time series data are abundant, as data are often generated over time in a way that introduces dependence between observations. 
However, the analysis of time series presents a unique challenge for both statisticians and practitioners, as the temporal dependence between observations requires careful treatment.
A unifying approach in time series analysis, known as state space modeling, assumes that observations from the system under investigation depend on an unobservable process.
Constructing a state space model involves proposing equations that describe the evolution of this unobservable process over time and the unobserved process is related to observations.
Common goals of state space modeling include making inferences about critical yet unobserved variables, incorporating real-world mechanisms into statistical models, and enabling likelihood based inference for models who that lack closed-form expressions of the likelihood function.

The term state space model (SSM) has a number of synonyms including mechanistic model, partially observed Markov process (POMP) model, and hidden Markov model (HMM) \citep{king15}.
In some literature, these terms have alternatively been used to refer to particular instances of the more general model class \citep[e.g.,][]{glennie23}, resulting in some confusion.
% To avoid ambiguity, unifying notation and terminology is presented in this section.
Throughout the dissertation, I follow the seminal work of \citet{durbin12} and use the term SSM to refer to the most general case.
That is, the observed time series is assumed to be reliant on an unobservable collection of variables at each observation time, without any restrictions on relationships between variables.
The term \emph{mechanistic} is reserved for models where the evolution of latent variables is dictated by equations intended to replicate real-world mechanisms, such as the transmission of infectious diseases among susceptible populations \citep[e.g.,][]{wheeler24}.
In most instances in this thesis and existing literature, the latent variables are treated as a Markov process, in which case the term POMP can be used.
Finally, existing literature often uses the term HMM to refer to POMP models where the latent variables take values in a discrete and finite space \citep{eddy04,doucet01,glennie23,newman23}; this same definition is used here.

\section{Defining State Space Models and Notation}\label{sec:intro-POMP}

In this section, some general terminology and notation that is common across chapters is introduced, and refined in subsequent chapters as needed. 
This thesis presents methodologies for modeling real-valued time series data, collected at observation times $t_1, t_2, \ldots, t_N \in \mathcal{T} \subset \R$.
Observation times do not need to be equally spaced.
Notably, this treatment includes time series where measurements are equally spaced but occasionally missing.

The observation at time $t_n$, $n \in \{1, \ldots, N\}$ can be vector valued, and is denoted $\bm{y}^*_{n} \in \R^{d_y}$.
The collection of observations $\{\bm{y}^*_{n},\, n\in 1, \ldots, N\}$ is modeled as a single realization of a collection of random variables $\{\bm{Y}_n, \, n\in 1, \ldots, N\}$, called the observable or measurement process.
The state space formulation introduces a collection of unobservable random variables $\{\bm{X}_n, \, n\in 1, \ldots, N\}$ that exist at the same time points as the observable process, and we generally assume that this unobservable (or latent) process takes values in a subset of $\R^{d_x}$.
Most often, models of interest include a description of how the latent process is initialized at some time $t_0 < t_1$, and the collection of latent variables is extended to include $\bm{X}_0$.
The latent process can be modeled either as a discrete or continuous time process.

For convenience, we adopt the notation that for any integers $a$ and $b$, $\seq{a}{b}$ is the vector $(a, a+1, \ldots, b-1, b)$, and use the convention that $\seq{a}{b} = \empty$ if $b < a$. 
Similarly, we write the entire collection of observations as $\bm{y}^*_{1:N} = (\bm{y}^*_1, \ldots, \bm{y}^*_N)$, and use the same basic notation for $Y_{1:N}$ and $X_{0:N}$.  
The joint probability density or (mass function) of the observable and unobservable processes is assumed to exist, and can be written as $f_{\bm{X}_{0:N}, \bm{Y}_{1:N}}(\bm{x}_{0:N}, \bm{y}_{1:N}; \, \paramVec)$, where $\paramVec$ is a parameter vector $\paramVec \in \R^{d_\paramVec}$.
The likelihood is a function of $\paramVec$, defined by the density of $\bm{Y}_{1:N}$ evaluated at the observed data $\bm{y}^*_{1:N}$, given by
\begin{eqnarray}
\label{eq:likedef}
\mathcal{L}(\paramVec) = f_{\bm{Y}_{1:N}}\big(\bm{y}_{1:N}^*; \, \paramVec\big) = \int f_{\bm{X}_{0:N}, \bm{Y}_{1:N}}\big(\bm{x}_{0:N}, \bm{y}_{1:N}^*;\, \paramVec\big) \, d\bm{x}_{0:N}.
\end{eqnarray}
In all but the simplest models, a closed-form expression of the likelihood function is not readily available due to the high-dimensional integral involving the latent variables.

The SSMs described in Chapters~\ref{chpt:arima}--\ref{chpt:mpif} can all be classified as POMP models.
This model class makes some additional, nonrestrictive assumptions that can be used to enable likelihood based inference.
The latent process is assumed to be Markovian and independent of the measurement process.
Additionally, the measurements are assumed to be conditionally independent given the current value of the latent process.
In terms of joint and conditional density functions of variables at observation times, these assumptions imply that for all $n \in \seq{1}{N}$,
$$
f_{\bm{X}_{n} | \bm{X}_{1:n-1}}(\bm{x}_{n} | \bm{x}_{1:n-1}; \, \paramVec) = f_{\bm{X}_{n} | \bm{X}_{n-1}}(\bm{x}_{n} | \bm{x}_{n-1}; \, \paramVec),
$$
and
$$
f_{\bm{Y}_{n} | \bm{X}_{1:N}, \bm{Y}_{-n}}(\bm{y}_{n} | \bm{x}_{0:N}, \bm{y}_{-n}; \, \paramVec) = f_{\bm{Y}_{n} | \bm{X}_{n}}(\bm{y}_{n} | \bm{x}_{n}; \, \paramVec),
$$
Using the notation that $\bm{Y}_{-n}$ is the collection of random variables $\bm{Y}_{1:N}$ without $\bm{Y}_{n}$.
A schematic diagram representing a general POMP model is given in Figure~\ref{fig:pompDiagram}
The additional assumptions of a POMP model can be used to refactor the joint density of latent and observable random variables as
\begin{eqnarray}
\label{eq:jointLik}
f_{\bm{X}_{0:N}, \bm{Y}_{1:N}}\big(\bm{x}_{0:N}, \bm{y}_{1:N};\, \paramVec\big) = f_{\bm{X}_0}\big(\bm{x}_0;\, \paramVec\big)\prod_{n = 1}^N f_{\bm{X}_n|\bm{X}_{n-1}}\big(\bm{x}_{n}|\bm{x}_{n-1}; \, \paramVec\big)f_{\bm{Y}_n|\bm{X}_{n}}\big(\bm{y}_n|\bm{x}_{n}; \, \paramVec\big).
\end{eqnarray}
This representation is useful as the entire model can be expressed in terms of three simple components, namely the \emph{initializer} $f_{\bm{X}_0}(\bm{x}_0; \, \paramVec)$, the \emph{process} or \emph{transition} model $f_{\bm{X}_{n} | \bm{X}_{n - 1}}(\bm{x}_n | \bm{x}_{n - 1}; \, \paramVec)$, and the \emph{measurement} model $f_{\bm{Y}_{n} | \bm{X}_n}(\bm{y}_n | \bm{x}_n ; \, \paramVec)$.
Each of these pieces can depend arbitrarily on observation time and the individual components of the parameter vector $\paramVec$.

\begin{figure}[!ht]
<<pompDiagram, echo=FALSE, fig.height=3, fig.width=6>>=
library(grid)
library(latex2exp)
vp <- viewport(x=unit(0.5,"npc"),y=unit(0.54,"npc"),
               width=unit(0.96,"npc"),height=unit(0.96,"npc"))
pushViewport(vp)

X_offset <- 0.04

fs <- 12
x1 <- 0.5+X_offset; y1 <- 0.88; offset <- 0.225;
gp <- gpar(lty=2,col=grey(0.6),fontsize=12)
grid.text(x=x1,y=y1,label="measurement model",just="centre",gp=gpar(lty=2,col=grey(0.6),fontsize=11))
grid.text(x=x1+offset,y=y1-0.005,TeX("$f_{Y_n|X_n}(y_n|x_n;\\,\\theta)$", italic = TRUE),gp=gpar(fontsize=11,col=grey(0.5)))
x1 <- (2* x1 + offset/2)/2+X_offset
grid.lines(x=unit(c(x1,7/24+X_offset),"npc")+unit(c(0,2),"points"),y=unit(c(y1,1/2),"npc")+unit(c(-fs/2,0),"points"),gp=gp)
grid.lines(x=unit(c(x1,11.8/24+X_offset),"npc")+unit(c(0,2),"points"),y=unit(c(y1,1/2),"npc")+unit(c(-fs/2,0),"points"),gp=gp)
grid.lines(x=unit(c(x1,18/24+X_offset),"npc")+unit(c(0,-2),"points"),y=unit(c(y1,1/2),"npc")+unit(c(-fs/2,0),"points"),gp=gp)

x1 <- 0.25+X_offset; y1 <- 0.14; offset <- 0.2075
grid.text(x=x1,y=y1,label="process model",just="centre",gp=gpar(lty=2,col=grey(0.6),fontsize=11))
grid.text(x=x1+offset,y=y1-0.005,TeX("$f_{X_n|X_{n-1}}(x_n|x_{n-1};\\,\\theta)$", italic = TRUE),gp=gpar(fontsize=11,col=grey(0.5)))
x1 <- (2 * x1 + offset) / 2
grid.lines(x=unit(c(x1,4/24+X_offset),"npc"),y=unit(c(y1,1/3),"npc")+unit(c((fs+1)/2,-fs/4),"points"),gp=gp)
grid.lines(x=unit(c(x1,9.5/24+X_offset),"npc"),y=unit(c(y1,1/3),"npc")+unit(c((fs+1)/2,-fs/4),"points"),gp=gp)
grid.lines(x=unit(c(x1,13.5/24+X_offset),"npc"),y=unit(c(y1,1/3),"npc")+unit(c((fs+1)/2,-fs/4),"points"),gp=gp)
grid.lines(x=unit(c(x1,19.6/24+X_offset),"npc"),y=unit(c(y1,1/3),"npc")+unit(c(fs/2,-fs/4),"points"),gp=gp)
grid.lines(x=unit(c(x1,16/24+X_offset),"npc"),y=unit(c(y1,1/3),"npc")+unit(c((fs+1)/2,-fs/4),"points"),gp=gp)

grid.text(x=1/72,y=c(1/3,2/3),label=c("latent states","observations"),just="centre",rot=90,gp=gpar(lty=2,col=grey(0.6),fontsize=11))

x1 <- unit(c(2, 7, 12, 18, 7, 12, 18)/24+X_offset, "npc")
y1 <- unit(c(rep(1,4),rep(2,3))/3,"npc")
w <- unit(1/12,"npc")
h <- unit(1/6,"npc")

grid.lines(x=c(1/48,47/48+X_offset/2),y=1/12,arrow=arrow(length=unit(0.02,"npc")))
grid.text(x=x1[1:4],y=1/24,label=c(expression(italic(t[0])),expression(italic(t[1])),expression(italic(t[2])),expression(t[n])))

grid.text(x=unit(15/24+X_offset, 'npc'), y=1/24,label=quote(phantom(0)~cdots~phantom(0)))
grid.text(x=unit(21/24+X_offset, 'npc'), y=1/24,label=quote(phantom(0)~cdots~phantom(0)))

grid.rect(x=x1,y=y1,width=w,height=h,just=c(0.5,0.5),gp=gpar(fill="white",lwd=2))
grid.text(x=x1,y=y1,label=c(
  expression(italic(X[0])),expression(italic(X[1])),
  expression(italic(X[2])),expression(italic(X[n])),
  expression(italic(Y[1])),
  expression(italic(Y[2])),expression(italic(Y[n]))),
  gp=gpar(fontface=3))
grid.text(x=c(15, 21)/24+X_offset,y=unit(1/3-(1/250),"npc")+unit(2,"point"),label=quote(phantom(0)~cdots~phantom(0)),gp=gpar(fontsize=15))
grid.lines(x=c(2,7)/24+c(1.2,-1.2)/24+X_offset,y=1/3,arrow=arrow(length=unit(0.02,"npc")),gp=gpar(lwd=2))
grid.lines(x=c(7,12)/24+c(1.2,-1.2)/24+X_offset,y=1/3,arrow=arrow(length=unit(0.02,"npc")),gp=gpar(lwd=2))
grid.lines(x=c(12,15)/24+c(1.2,-0.7)/24+X_offset,y=1/3,arrow=arrow(length=unit(0.02,"npc")),gp=gpar(lwd=2))
grid.lines(x=c(15,18)/24+c(0.7,-1.2)/24+X_offset,y=1/3,arrow=arrow(length=unit(0.02,"npc")),gp=gpar(lwd=2))
grid.lines(x=c(18,21)/24+c(1.2,-0.6)/24+X_offset,y=1/3,arrow=arrow(length=unit(0.02,"npc")),gp=gpar(lwd=2))

grid.lines(x=7/24+X_offset,y=c(1,2)/3+c(1,-1)/12,arrow=arrow(length=unit(0.02,"npc")),gp=gpar(lwd=2))
grid.lines(x=12/24+X_offset,y=c(1,2)/3+c(1,-1)/12,arrow=arrow(length=unit(0.02,"npc")),gp=gpar(lwd=2))
grid.lines(x=18/24+X_offset,y=c(1,2)/3+c(1,-1)/12,arrow=arrow(length=unit(0.02,"npc")),gp=gpar(lwd=2))

popViewport()
@
\caption{\label{fig:pompDiagram}A flow diagram representing an arbitrary POMP model.}
\end{figure}

Though these additional assumptions do not generally give rise to a closed-form expression of the likelihood function, they do enable likelihood based inference in many scenarios.
For instance, if the evolution of the latent variables and measurements are dictated by linear Gaussian equations, then the Kalman filter \citep{kalman60} can be used to evaluate the likelihood function exactly.
These types of models are generally referred to as linear Gaussian SSMs rather than linear Gaussian POMP models, and I use this terminology as well in this thesis when referring to this model class.
Despite the ability to evaluate the likelihood function exactly, linear Gaussian SSMs can still result in challenges for inference \citep{auger16}.
This is demonstrated in Chapter~\ref{chpt:arima}, where I demonstrate that frequently used algorithms for maximizing the likelihood of a specific class of linear Gaussian SSMs often result in sub optimal parameter estimates, and propose a novel algorithm (Algorithm~\ref{alg:mle}) that addresses these shortcomings.

Likelihood-based inference for POMP models has often been successful in more general cases, particularly when the observed variables are univariate ($d_y = 1$) or when the dimensionality is low ($d_y < 5$).
Although the likelihood function cannot be evaluated exactly, various algorithms can approximate it using Monte Carlo techniques.
For example, in mechanistic models of infectious disease dynamics, the particle filter \citep{arulampalam02} is commonly used to evaluate model likelihoods. 
Additionally, iterated filtering, an extension of the particle filter, is used to maximize the likelihood \citep{ionides15}.
An application of this approach is demonstrated using Model~1 of Chapter~\ref{chpt:haiti}. 
However, the the applicability of these Monte Carlo approaches diminishes quickly as the dimensionality of the model grows.

A common instance that has challenged existing Monte Carlo procedures arises when a collection of similar but independent POMP models is used to analyze a set of time series, known as panel data or longitudinal data.
This assembly of POMP models is referred to as a PanelPOMP model.
Inference methodologies for this class of models is the primary focus of Chapter~\ref{chpt:mpif}.
The independence of the models allows the particle filter to be applied independently to each model for likelihood evaluation.
However, when parameters are shared across models, coupling is introduced, which can lead to scaling issues typical of Monte Carlo algorithms in high-dimensional settings.
The MPIF algorithm, introduced in Chapter~\ref{chpt:mpif}, mitigates these issues and significantly enhances the ability to fit PanelPOMPs to panel data compared to existing approaches.

High-dimensional models can also emerge when time series data are collected from multiple sites or individuals and the latent processes shares dynamics across measurement locations.
This situation commonly arises when there is spatial dependence between measurement locations, giving rise to the name SpatPOMP for this class of models.
Although the primary focus of this thesis is not on methods for SpatPOMP models, Chapter~\ref{chpt:haiti} explores an application of a SpatPOMP model for cholera infections in the 2010-2019 outbreak in Haiti.
The model accounts for spatial dependence between infections recorded at distinct Haitian administrative d\'{e}partements.

\section{Overview of remaining chapters}\label{intro:overview}

Chapter~\ref{chpt:arima} describes an algorithm for performing maximum likelihood estimation for parameters of arbitrary autoregressive moving average (ARMA) models.
Existing methodologies for this task express the ARMA model as a linear Gaussian SSM, and then use the Kalman filter \citep{kalman60} to evaluate the likelihood function;
the likelihood is then maximized using numeric optimization algorithms.
In this chapter, I demonstrate that this approach often results in sub optimal parameter estimates, as numeric optimizers frequently converge to local maxima of the likelihood surface.
To overcome these optimization deficiencies, I propose a random parameter initialization algorithm that enables sampling of ARMA parameter coefficients, accounting for the complex geometry of the parameter space.
Combined with existing methodologies for fitting ARMA parameters, this new approach is demonstrated to frequently result in improved parameter estimates using a large number of simulated data.

In Chapter~\ref{chpt:haiti}, I examine the role SSMs have in influencing policy decisions regarding infectious disease outbreaks.
This is done by reviewing the statistical models that were used to analyze the 2019-2019 cholera outbreak in Haiti.
Following this review, a retrospective analysis of the cholera outbreak is conducted using three previously proposed SSMs.
Using the lessons learned by reviewing existing approaches, the retrospective analysis results in a reproducible workflow for developing models, evaluating model quality, and refining the resulting models in order to improve statistical outcomes.
This project also demonstrates the efficacy of recently developed methodology for likelihood based inference of SpatPOMP models.
This work has been been published in PLOS Computational Biology \citep{wheeler24}.

Chapter~\ref{chpt:mpif} proposes a simulation-based algorithm designed to perform maximum likelihood estimation for a class of high-dimensional POMP models.
This algorithm, called the Marginalized Panel Iterated Filter (MPIF), significantly enhances the capability of iterated filtering algorithms to estimate parameters for large collections of independent but related POMP models.
POMP models are routinely fitted to univariate time series data, but seldom to collections of time series---called panel data.
This suggests that there are practical difficulties in using existing approaches for this task.
The MPIF algorithm improves empirical convergence rates of existing procedures by addressing the issue of particle depletion that occurs when performing iterated filtering on models that have high-dimensional parameter spaces.
Theoretical support for the algorithm is provided through an analysis of iterating marginalized Bayes maps.
Additionally, asymptotic theory demonstrating the convergence of general iterated filtering algorithms for PanelPOMP models without the marginalization step is presented.
This work is currently being prepared for submission, and reflects joint work with my adviser and fellow Ph.D. student, Aaron Abkemeier.

The final chapter provides a brief summary of the work presented in Chapters~\ref{chpt:arima}--\ref{chpt:mpif}.
This is followed by a discussion on future work related to this thesis.